---
title: "thesis_code"
output: pdf_document
---

## Refactoring


#### Generating data
Idea: start at a random point, choose an initial direction, move in that direction continuously, change direction (within 15 degrees) with prob=p, change direction more (within 30 degrees) with prob=q, and so on; keep track of truth and add noise as you go. 
```{r}
generate_filament <- function(n, step_size, bool_func) {
  # bool_func: generate function by restricting theta
  x <- c(); y <- c()
  x[1] <- runif(1)
  y[1] <- runif(1)
  theta <- runif(1, 0, 360)
  if (bool_func) {
    theta <- runif(1, -30, 30)
  }
  
  data.x <- c()
  data.y <- c()
  
  for (i in 2:n) {
    # idea 1: generate data on axis orthogonal to direction
    # idea 2: generate data in bivariate normal around point
    
    perp_theta <- theta + 90
    # should implement random number of points
    # num_points <- runif(1, 0, 2)
    dist <- rnorm(1, 0, 0.08)
    data.x[i-1] <- x[i-1] + dist*cos(perp_theta*pi/180)
    data.y[i-1] <- y[i-1] + dist*sin(perp_theta*pi/180)
    
    x[i] <- x[i-1] + step_size*cos(theta*pi/180)
    y[i] <- y[i-1] + step_size*sin(theta*pi/180)
    
    if (bool_func) {
      old_theta <- theta
      theta <- theta + rnorm(1, 0, 15)
      while (theta > 90 | theta < -90) {
        theta <- old_theta + rnorm(1, 0, 15)
      }
    }
    else { theta <- theta + rnorm(1, 0, 15) }
    # could implement some idea of inertia to prevent sharp switchbacks/encourage smoothness?
  }
  
  # scale to [0,1]
  #x_scale <- max(abs(x))
  #y_scale <- max(abs(y))
  
  #data.x < data.x/x_scale
  #data.y <- data.y/y_scale
  #x <- x/x_scale
  #y <- y/y_scale
  
  #plot(x, y, xlim = 0:1, ylim = 0:1, type='l')
  x <- (x-min(x))/(max(x)-min(x)) 
  y <- (y-min(y))/(max(y)-min(y))
  data.x <- (data.x-min(data.x))/(max(data.x)-min(data.x))
  data.y <- (data.y-min(data.y))/(max(data.y)-min(data.y))
  #plot(x, y, type='l')
  #points(data.x, data.y)
  return(cbind(data.x, data.y))
}

plot(generate_filament(100, .05, TRUE))
```


#### Helper Functions

to_vec: returns line segment
```{r}
library(zeallot)
# params: center, angle (rad), magnitude
# return: vector
to_vec <- function(x, y, magnitude, theta) {
  return(c(cos(theta)*magnitude, sin(theta)*magnitude))
}
```


Constrained mean-shift algorithm

```{r}
bi_normal_kernel_y <- function(x, y, data, h) {
  # data is nx2 matrix
  x_data <- data[,1]; y_data <- data[,2]
  # what if you wanted both x and y? would there be an easier way that just returning both?
  return(sum((dnorm((x-x_data)/h))*(dnorm((y-y_data)/h))*y_data)/sum((dnorm((x-x_data)/h))*(dnorm((y-y_data)/h))))
}
```

Adapt mean shift to generalized direction

```{r}
constrained_mean_shift <- function(pt, theta, data, h) {
  x <- pt[1]; y <- pt[2]
  #x_data <- data[,1]; y_data <- data[,2]
  
  # change of basis matrix
  cob <- rbind(c(cos(theta), -sin(theta)), c(sin(theta), cos(theta)))
  
  # right multiply to transform
  new_data <- data %*% cob
  new_x <- new_data[,1]
  new_y <- new_data[,2]
  
  rotated_pt <- pt %*% cob
  
  # kernel weighting
  kw <- dnorm((x-new_x)/h)#*dnorm((y-new_y)/h)
  new_pt <- c(rotated_pt[1], sum(new_y*kw)/sum(kw))
  return(new_pt %*% t(cob))
}
```


Returns logical vector of points in ellipse
```{r}
in_ellipse <- function(data, center, a, b, theta) {
# params:
  # point (x, y)
  # center (x, y)
  # a (major axis)
  # b (minor axis)
  # theta (angle of rotation)
  
  #plot(data, asp = 1)
  #draw.ellipse(center[1], center[2], a=a, b=b, angle=theta, deg=FALSE)
  
  pt_in_ellipse <- c()
  
  for (i in 1:nrow(data)) {
    centered_point <- data[i,]-center
    x <- centered_point[1]; y <- centered_point[2]
    pt_in_ellipse[i] <- ((x*cos(theta)+y*sin(theta))^2/a^2 + (x*sin(theta)-y*cos(theta))^2/b^2) <= 1
  }
  return(pt_in_ellipse)
}

#in_ellipse(cbind(c(1, 2, 3), c(1, 2, 3)), c(2, 2), .5, .5, 0)
```

Choose random start points, excluding points within neighborhoods of those already chosen

```{r}
get_centers <- function(data, r) {
  centers <- c()
  bool <- logical(nrow(data))
  while (!all(bool)) {
    center <- sample(which(!bool), 1)
    #while (bool[center] == FALSE) center <- sample(nrow(data), 1)
    pt <- data[center,]
    centers <- c(centers, center)
    bool[which(in_ellipse(data, pt, r, r, 0))] <- TRUE
  }
  return(centers) # returns as index of input data
}
```

smooth: 
data params x, y
bandwith param bw

returns smoothed estimate as n x 2 array

```{r}
# returns n x 2 matrix of x and y values
smooth <- function(x, y, bw) {
  n <- length(x)
  smooth.x <- c()
  smooth.y <- c()
  for (i in 1:n) {
    if (i-bw < 1) {
      bw1 <- 1
      bw2 <- 2*i-1
    }
    else if (i+bw > n) {
      bw1 <- i-(n-i)
      bw2 <- n
    }
    else {
      bw1 <- i-bw
      bw2 <- i+bw
    }
    #bw1 <- max(1, i-bw)  #need to force a symmetric kernel here -> if i-bw = 1, bw2 -> i+1
    #bw2 <- min(i + bw, length(new.x))
    smooth.x[i] <- mean(x[bw1:bw2], na.rm=TRUE)
    smooth.y[i] <- mean(y[bw1:bw2], na.rm=TRUE)
  }
  return(cbind(smooth.x, smooth.y))
}
```

generate_grid: 
data params x, y
grid size param eps

returns gridding along input points as n x 2 array

```{r}
# input: grid ordered by x-values
generate_grid <- function(x, y, eps) {
  d <- c()
  for (i in 1:(length(x)-1)) {
   d[i] <- sqrt((x[i+1]-x[i])^2+(y[i+1]-y[i])^2)
  }
  lastx <- 1
  new.x <- x[1]
  new.y <- y[1]
  k <- 2
  while (lastx < length(x)) {
    if (d[lastx] > eps) {
      weight <- eps/d[lastx]
      new.x[k] <- x[lastx+1]*weight + (1-weight)*x[lastx]
      new.y[k] <- y[lastx+1]*weight + (1-weight)*y[lastx]
      d[lastx] <- d[lastx]-eps
      x[lastx] <- new.x[k]
      y[lastx] <- new.y[k]
      k <- k + 1
    } else {
      eps1 <- eps-d[lastx]
      lastx <- lastx + 1
      weight <- eps1/d[lastx]
      new.x[k] <- x[lastx+1]*weight + (1-weight)*x[lastx]
      new.y[k] <- y[lastx+1]*weight + (1-weight)*y[lastx]
      d[lastx] <- d[lastx]-eps1
      x[lastx] <- new.x[k]
      y[lastx] <- new.y[k]
      k <- k + 1
    }
  }
  new.x[k] <- x[length(x)]
  new.y[k] <- y[length(y)]
  return(cbind(na.omit(new.x), na.omit(new.y)))
}

# want to return:
  # new x, new y
  # direction to next x, y
  # perpendicular direction
```



```{r}
x <- seq(0, 2, by=0.03)
y <- sin(x*3)/2
plot(x, y, cex=0.1)

grid <- generate_grid(x, y, 0.01)
new.x <- na.omit(grid[,1])
new.y <- na.omit(grid[,2])

d1 <- c()
for (i in 1:(length(new.x)-1)) {
 d1[i] <- sqrt((new.x[i+1]-new.x[i])^2+(new.y[i+1]-new.y[i])^2)
} 

#   w <- c(x[i+1]-x[i], y[i+1]-y[i])
#   w <- w/sqrt(sum(w^2))
#   x[i+1] <- x[i] + w[1]*eps
#   y[i+1] <- y[i] + w[2]*eps
# }
#points(new.x, new.y, col='red', cex=.1)
bw <- 25
smooth.x <- smooth(new.x, new.y, bw)[,1]
smooth.y <- smooth(new.x, new.y, bw)[,2]
lines(smooth.x, smooth.y)

smooth.grid <- generate_grid(smooth.x, smooth.y, 0.008)
smooth.grid.x <- smooth.grid[,1]
smooth.grid.y <- smooth.grid[,2]

points(smooth.grid.x, smooth.grid.y, col='blue', cex=.1)

d1 <- c()
for (i in 1:(length(smooth.grid.x)-1)) {
  d1[i] <- sqrt((smooth.grid.x[i+1]-smooth.grid.x[i])^2+(smooth.grid.y[i+1]-smooth.grid.y[i])^2)
} 
```

mapping: figure out which grid points it's closest to, then project onto the line to find mapping


```{r}
x_data <- seq(0, 2, by=0.03)
y_data <- sin(x_data*3)/2 + rnorm(length(x_data), 0, 0.05)
plot(smooth.grid.x, smooth.grid.y, cex=0.1)
points(x_data, y_data)

data <- cbind(x_data, y_data)
grid <- cbind(smooth.grid.x, smooth.grid.y)

pt <- data[30,]

pts <- data[c(6, 19, 30, 36, 48),]
points(pts[,1], pts[,2], col='red')
points(pt[1], pt[2], col="red")

# append data point to grid matrix
grid_with_point <- rbind(pt, grid)
dists <- as.matrix(dist(grid_with_point))[-1,1] # should correct the indices here to avoid confusion
t <- which.min(dists)
if (dists[t+1] < dists[t-1]) {
  a_index <- t
  b_index <- t+1
} else {
  a_index <- t-1
  b_index <- t
}

# get a, b
a <- grid[a_index,]
b <- grid[b_index,]

# projection of v onto h (horizontal direction)
v <- pt-a
h <- b-a

proj <- (sum(v*h)/sum(h*h))*h
x_dist <- sqrt(sum(proj*proj))

# v minus the above projection (vertical direction)
y_proj <- v-proj
y_dist <- sqrt(sum(y_proj*y_proj))
if (y_proj[2] < 0) y_dist <- y_dist*(-1)
```

get_mapping:
data param data
grid 

```{r}
# find nearest by taking min of distance from all grid points, then just need to determine if right or left

get_mapping <- function(data, grid) {
  mapped_data_x <- c()
  mapped_data_y <- c()
  
  d <- as.matrix(dist(grid))
  map <- c(0, cumsum(d[row(d)+1 == col(d)]))
  
  for (i in 1:nrow(data)) {
    pt <- data[i,]
    grid_with_point <- rbind(pt, grid)
    dists <- as.matrix(dist(grid_with_point))[-1,1] # should correct the indices here to avoid confusion
    t <- which.min(dists)
    if (t == 1) {
      a_index <- t
      b_index <- t+1
    } else if (t == length(dists)) {
      a_index <- t-1
      b_index <- t
    } else if (dists[t+1] < dists[t-1]) {
      a_index <- t
      b_index <- t+1
    } else {
      a_index <- t-1
      b_index <- t
    }
    
    # get a, b
    a <- grid[a_index,]
    b <- grid[b_index,]
    
    # projection of v onto h (horizontal direction)
    v <- pt-a
    h <- b-a
    
    proj <- (sum(v*h)/sum(h*h))*h
    x_dist <- sqrt(sum(proj*proj))
    
    # v minus the above projection (vertical direction)
    y_proj <- v-proj
    y_dist <- sqrt(sum(y_proj*y_proj))
    
    cross_prod <- v[1]*h[2]-h[1]*v[2]
    y_dist <- y_dist - 2*y_dist*(cross_prod > 0)
    #if (cross_prod < 0) y_dist <- y_dist*(-1)
    
    #arcl <- t*eps
    # 0.08 is not exact -- need to find by taking distances from distance matrix FIX THIS
    mapped_data_x[i] <- map[a_index]+x_dist 
    mapped_data_y[i] <- y_dist
  }
  return(cbind(mapped_data_x, mapped_data_y))
}

mapped_data <- get_mapping(data, grid)
plot(data, col=1+(mapped_data[,2]>0))
points(grid, col='blue', cex=0.2)


mapped_data <- get_mapping(data, grid)
plot(mapped_data[,1], mapped_data[,2])
mapped_pts <- mapped_data[c(6, 19, 30, 36, 48),]
points(mapped_pts[,1], mapped_pts[,2], col='red')
abline(h=0)
```

```{r}
# remember mapping distances
d <- as.matrix(dist(grid))
map <- c(0, cumsum(d[row(d)+1 == col(d)]))

# recover mapping directions
dirs <- c()
for (i in 1:length(map)-1) {
  v <- grid[i+1,]-grid[i,]
  dirs <- rbind(dirs, v/sqrt(sum(v^2)))
}

inverted_data <- c()

for (i in 1:nrow(mapped_data)) {
  x <- mapped_data[i,1]
  y <- mapped_data[i,2]
  
  # which grid index corresponds to the point
  grid_ind <- which.min(x-map + 100*((x-map) < 0))
  x_pos <- (x-map[grid_ind])
  y_pos <- y
  
  # in the original space
  d <- dirs[grid_ind,]
  d_ortho <- c(d[1], -d[2])
  
  # theta
  theta <- atan2(y_pos, x_pos)
  rotation <- rbind(c(cos(theta), -sin(theta)), c(sin(theta), cos(theta)))
  v <- d*sqrt(x_pos^2+y_pos^2)
  new_v <- as.vector(rotation %*% v)
  
  #inverted_data <- rbind(inverted_data, grid[grid_ind,] + x_pos*d + y_pos*d_ortho)
  inverted_data <- rbind(inverted_data, grid[grid_ind,] + new_v)
}

plot(inverted_data[,1], inverted_data[,2])
points(data, col='blue', cex=0.5)
sum((inverted_data-data)^2)
```

Invert mapping function

```{r}
get_invert_mapping <- function(mapped_data, grid) {
  # remember mapping distances
  d <- as.matrix(dist(grid))
  map <- c(0, cumsum(d[row(d)+1 == col(d)]))
  
  # recover mapping directions
  dirs <- c()
  for (i in 1:length(map)-1) {
    v <- grid[i+1,]-grid[i,]
    dirs <- rbind(dirs, v/sqrt(sum(v^2)))
  }
  dirs <- rbind(dirs, dirs[(length(map)-1),]) #workaround until I can fix this issue
  
  inverted_data <- c()
  
  for (i in 1:nrow(mapped_data)) {
    x <- mapped_data[i,1]
    y <- mapped_data[i,2]
    
    # which grid index corresponds to the point
    grid_ind <- which.min(x-map + 100*((x-map) < 0))
    x_pos <- (x-map[grid_ind])
    y_pos <- y
    
    # in the original space
    d <- dirs[grid_ind,]
    d_ortho <- c(d[1], -d[2])
    
    # theta
    theta <- atan2(y_pos, x_pos)
    rotation <- rbind(c(cos(theta), -sin(theta)), c(sin(theta), cos(theta)))
    v <- d*sqrt(x_pos^2+y_pos^2)
    new_v <- as.vector(rotation %*% v)
    
    #inverted_data <- rbind(inverted_data, grid[grid_ind,] + x_pos*d + y_pos*d_ortho)
    inverted_data <- rbind(inverted_data, grid[grid_ind,] + new_v)
  }
  
  return(inverted_data)
}
```


```{r}
inverted_data <- get_invert_mapping(mapped_data, grid)
plot(inverted_data)
points(data, cex=0.5)
```



```{r}
library(zeallot)
library(plotrix)
# choose initial point in warped space
pt <- sample(1:nrow(mapped_data), 1)
plot(mapped_data)
c(x, y, a, b, theta) %<-% local_pca_expand(pt, mapped_data, .3, 3)
```


```{r}
generate_spiral <- function(n, step_size) {
  x <- c(); y <- c()
  x[1] <- 0
  y[1] <- 0
  theta <- 0
  r <- 0
  
  data.x <- c()
  data.y <- c()
  
  for (i in 2:n) {
    # idea 1: generate data on axis orthogonal to direction
    # idea 2: generate data in bivariate normal around point
    
    x[i] <- r*cos(theta*pi/180)
    y[i] <- r*sin(theta*pi/180)
    
    theta <- theta + 5
    r <- r + 0.01
  }
  return(cbind(x, y))
}

plot(generate_spiral(120, 0.05))
```

```{r}
generate_circle <- function(n, step_size) {
  # bool_func: generate function by restricting theta
  x <- c(); y <- c()
  x[1] <- runif(1)
  y[1] <- runif(1)
  theta <- runif(1, 0, 360)
  
  data.x <- c()
  data.y <- c()
  
  for (i in 2:n) {
    # idea 1: generate data on axis orthogonal to direction
    # idea 2: generate data in bivariate normal around point
    
    perp_theta <- theta + 90
    # should implement random number of points
    # num_points <- runif(1, 0, 2)
    dist <- rnorm(1, 0, 0.08)
    data.x[i-1] <- x[i-1] + dist*cos(perp_theta*pi/180)
    data.y[i-1] <- y[i-1] + dist*sin(perp_theta*pi/180)
    
    
    x[i] <- x[i-1] + step_size*cos(theta*pi/180)
    y[i] <- y[i-1] + step_size*sin(theta*pi/180)
    
    theta <- theta + 3
  }
  
  x <- (x-min(x))/(max(x)-min(x)) 
  y <- (y-min(y))/(max(y)-min(y))
  data.x <- (data.x-min(data.x))/(max(data.x)-min(data.x))
  data.y <- (data.y-min(data.y))/(max(data.y)-min(data.y))
  
  return(cbind(data.x, data.y))
}

plot(generate_spiral(120, 0.05))
```

```{r}
add_noise <- function(data) {
  min.x <- min(data[,1])
  max.x <- max(data[,1])
  min.y <- min(data[,2])
  max.y <- max(data[,2])
  
  noise.x <- runif(nrow(data)/2, min.x, max.x)
  noise.y <- runif(nrow(data)/2, min.y, max.y)
  noise <- cbind(noise.x, noise.y)
  return(rbind(data, noise))
}
```

```{r}
get_pts_on_line <- function(data, centers, eps) {
  points <- c()
  prev <- data[centers[1],]
  for (i in 2:length(centers)) {
    curr <- data[centers[i],]
    v <- curr-prev
    v_norm <- sqrt(sum(v^2))
    v <- v/v_norm
    d <- 0
    while(d < v_norm) {
      points <- rbind(points, prev+d*v)
      d <- d + eps
    }
    prev <- curr
  }
  return(points)
}
```

```{r}
get_pts_on_line2 <- function(data, pts, eps) {
  line <- c()
  prev <- pts[1,]
  for (i in 2:nrow(pts)) {
    curr <- pts[i,]
    v <- curr-prev
    v_norm <- sqrt(sum(v^2))
    v <- v/v_norm
    d <- 0
    while(d < v_norm) {
      line <- rbind(line, prev+d*v)
      d <- d + eps
    }
    prev <- curr
  }
  return(line)
}
```

Test on generated data

```{r}
set.seed(47)
data <- generate_filament(100, 0.05, FALSE)
#data <- add_noise(generate_circle(120, 0.05))
centers <- c(1, sort(get_centers(data, 0.2)), nrow(data))
plot(data)
```
```{r}
centers <- sort(get_centers(data, 0.2))
new_centers <- c()
for (i in 1:length(centers)) {
  pts <- data[in_ellipse(data, data[centers[i],], 0.2, 0.2, 0),]
  if (length(pts) <= 2) {
    next
  }
  pca <- prcomp(pts)
  theta <- atan2(pca$rotation[2,1], pca$rotation[1,1])
  new_centers <- rbind(new_centers, constrained_mean_shift(as.vector(data[centers[i],]), theta, pts, 0.2))
}
plot(data)
points(data[centers,], col='blue')
points(new_centers, col='red')
```


```{r}
plot(data)
#line <- get_pts_on_line(data, centers, 0.01)
line <- get_pts_on_line2(data, new_centers, 0.01)
points(line, cex=.1)
```


```{r}
new.x <- line[,1]
new.y <- line[,2]

bw <- 10
smooth.x <- smooth(new.x, new.y, bw)[,1]
smooth.y <- smooth(new.x, new.y, bw)[,2]
plot(data)
lines(smooth.x, smooth.y)
```

```{r}
smooth.x <- inverted_data[,1]
smooth.y <- inverted_data[,2]
```


```{r}
smooth.grid <- generate_grid(smooth.x, smooth.y)
smooth.grid.x <- smooth.grid[,1]
smooth.grid.y <- smooth.grid[,2]

plot(data)
points(smooth.grid, col='blue', cex=.1)
```

```{r}
warped_data <- get_mapping(data, smooth.grid)
plot(warped_data)
abline(h=0, col='blue')
```

```{r}
warped_centers <- sort(get_centers(warped_data, 0.2))

new_warped_centers <- c()
for (i in 1:length(warped_centers)) {
  pts <- warped_data[in_ellipse(warped_data, warped_data[warped_centers[i],], 0.05, 0.05, 0),]
  if (length(pts) <= 2) {
    next
  }
  pca <- prcomp(pts)
  theta <- atan2(pca$rotation[2,1], pca$rotation[1,1])
  new_warped_centers <- rbind(new_warped_centers, constrained_mean_shift(as.vector(warped_data[warped_centers[i],]), theta, pts, 0.05))
}

warped_line <- get_pts_on_line2(warped_data, new_warped_centers, 0.01)

plot(warped_data)
#points(warped_data[warped_centers,], col='blue')
points(new_warped_centers, col='red')
lines(warped_line)
```
```{r}
warped.x <- warped_line[,1]
warped.y <- warped_line[,2]

bw <- 10
smooth.warped.x <- smooth(warped.x, warped.y, bw)[,1]
smooth.warped.y <- smooth(warped.x, warped.y, bw)[,2]

smooth_warped_line <- cbind(smooth.warped.x, smooth.warped.y)
plot(warped_data)
lines(smooth_warped_line)
```


JUST A TEST
```{r}
d <- as.matrix(dist(smooth.grid))
map <- c(0, cumsum(d[row(d)+1 == col(d)]))

# recover mapping directions
dirs <- c()
for (i in 1:length(map)-1) {
  v <- smooth.grid[i+1,]-smooth.grid[i,]
  dirs <- rbind(dirs, v/sqrt(sum(v^2)))
}
dirs <- rbind(dirs, dirs[(length(map)-1),])
```

```{r}
inverted_data <- get_invert_mapping(smooth_warped_line, smooth.grid)
plot(data)
#points(smooth.grid, cex=0.5, col='blue')
lines(inverted_data, col='blue')
```
```{r}
line <- inverted_data
```


```{r}
iterate <- function(data, estimate) {
  x <- estimate[,1]
  y <- estimate[,2]
  
  bw <- 8
  smooth.x <- smooth(x, y, bw)[,1]
  smooth.y <- smooth(x, y, bw)[,2]
  
  grid <- generate_grid(smooth.x, smooth.y, 0.008)
  
  warped_data <- get_mapping(data, grid)
  warped_centers <- c(1, sort(get_centers(warped_data, 0.2)), nrow(warped_data))
  warped_line <- get_pts_on_line(warped_data, centers, 0.01)
  #warped_line <- smooth(warped_line[,1], warped_line[,2], bw)
  
  inverted_estimate <- get_invert_mapping(warped_line, grid)
  return(inverted_estimate)
}

estimate <- line
for (i in 1:5) {
  estimate <- iterate(data, estimate)
}
plot(data)
#points(iterate(data, line), cex=0.5, col='red')
smooth_estimate <- smooth(estimate[,1], estimate[,2], 8)
lines(smooth_estimate, cex=0.5, col='red')

plot(get_mapping(data, estimate))
```

```{r}
iterate2 <- function(data, estimate) {
  x <- estimate[,1]
  y <- estimate[,2]
  
  grid <- generate_grid(x, y, 0.008)
  
  warped_data <- get_mapping(data, grid)
  warped_centers <- c(1, sort(get_centers(warped_data, 0.3)), nrow(warped_data))
  #warped_line <- get_pts_on_line(warped_data, centers, 0.01)
  #warped_line <- smooth(warped_line[,1], warped_line[,2], bw)
  
  #warped_centers <- sort(get_centers(warped_data, 0.2))

  new_warped_centers <- c()
  for (i in 1:length(warped_centers)) {
    pts <- warped_data[in_ellipse(warped_data, warped_data[warped_centers[i],], 0.05, 0.05, 0),]
    if (length(pts) <= 2) {
      next
    }
    pca <- prcomp(pts)
    theta <- atan2(pca$rotation[2,1], pca$rotation[1,1])
    new_center <- constrained_mean_shift(as.vector(warped_data[warped_centers[i],]), theta, pts, 0.05)
    print(new_center)
    if (is.na(new_center)) new_center <- as.vector(warped_data[warped_centers[i],])
    new_warped_centers <- rbind(new_warped_centers, new_center)
  }  # do this until convergence
  
  
  # what's causing N/A values??
  
  #new_warped_centers <- na.omit(new_warped_centers)
  
  warped_line <- get_pts_on_line2(warped_data, new_warped_centers, 0.01)
  
  warped.x <- warped_line[,1]
  warped.y <- warped_line[,2]
  
  bw <- 10
  
  smooth_warped_line <- smooth(warped.x, warped.y, bw)
  
  inverted_estimate <- get_invert_mapping(smooth_warped_line, grid)
  return(inverted_estimate)
}

set.seed(22)
data <- generate_filament(100, 0.05, TRUE)
plot(data)

centers <- sort(get_centers(data, 0.2))
new_centers <- c()
for (i in 1:length(centers)) {
  pts <- data[in_ellipse(data, data[centers[i],], 0.2, 0.2, 0),]
  if (length(pts) <= 2) {
    next
  }
  pca <- prcomp(pts)
  print(pca)
  theta <- atan2(pca$rotation[2,1], pca$rotation[1,1])
  new_centers <- rbind(new_centers, constrained_mean_shift(as.vector(data[centers[i],]), theta, pts, 0.2))
}


plot(data)
points(data[centers,], col='blue')
points(new_centers, col='red')

plot(data)
#line <- get_pts_on_line(data, centers, 0.01)
line <- get_pts_on_line2(data, new_centers, 0.01)
points(line, cex=.1)

estimate <- line
for (i in 1:6) {
  print(i)
  estimate <- iterate2(data, estimate)
}
plot(data)
#points(new_centers, col='blue')
points(estimate, cex=0.5, col='red')
#points(iterate2(data, line), cex=0.5, col='red')
smooth_estimate <- smooth(estimate[,1], estimate[,2], 8)
lines(smooth_estimate, cex=0.5, col='red')

plot(get_mapping(data, estimate))
```

1. use arias castro to find the ordering of the connecting points
2. mean shift y in warped space at 0, arc length to prevent shrinkage



```{r}
get_pts_on_line2 <- function(data, pts, eps) {
  line <- c()
  prev <- pts[1,]
  for (i in 2:nrow(pts)) {
    curr <- pts[i,]
    #print(curr)
    v <- curr-prev
    v_norm <- sqrt(sum(v^2))
    v <- v/v_norm
    d <- 0
    while(d < v_norm) {
      line <- rbind(line, prev+d*v)
      d <- d + eps
    }
    prev <- curr
  }
  return(line)
}
```


```{r}
library(zeallot)
library(plotrix)
# choose initial point in warped space
pt <- sample(1:nrow(warped_data), 1)
plot(warped_data)
c(x, y, a, b, theta) %<-% local_pca_expand(pt, warped_data, .3, 3)
plot(warped_data)

get_centers_pca <- function(data, r) {
  centers <- c()
  bool <- logical(nrow(data))
  while (!all(bool)) {
    center <- sample(which(!bool), 1)
    pt <- data[center,]
    centers <- c(centers, center)
    c(x, y, a, b, theta) %<-% local_pca_expand(center, data, r, 3)
    #print(c(x, y, a, b, theta))
    bool[which(in_ellipse(data, pt, a, b, theta))] <- TRUE
    print(bool)
  }
  return(centers) # returns as index of input data
}

centers <- c(1, sort(get_centers_pca(warped_data, .3)), nrow(warped_data))
```